---
title: "Classwork Worksheet 1"
output: html_document
author: TASHAE DEWALT
date: "2026-07-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Questions

1.What dataset are you working with for your final and provide a brief synopsis of what it contains? 

```{r}
#Art History
#Assess the demographic representation of artists through two popular textbooks and museums
```

2. What is a research question of interest to you relating to this data? Provide the question in a PICO format: Problem/Population Intervention/Exposure Comparison Outcome. https://pmc-ncbi-nlm-nih-gov.ezproxy.med.nyu.edu/articles/PMC11129835/ 

```{r}
#Probelm: The lack of visibility and representation transcends medicine, this can increases about minority populations
#Intervention: Using an art dataset for easier relatibility of subject, will then discuss similiarities of this to the world of research, science, and medicine
#Comparison: The exposure of differen ethnic backgroups in a medium outside of medicine, show how they are connected or how one can impact the other
#Outcome: Build upon this dataset for possible STEAM students/ people (Me), where the worlds may not converse much, this is a great opportunity to encourage both and increase inclusivity and trust of minority populations, especially here in America
```


3. Does this question meet the FINER criteria? Break down letter by letter to demonstrate. https://pmc-ncbi-nlm-nih-gov.ezproxy.med.nyu.edu/articles/PMC11129835/

```{r}
# Feasible? yes Interesting? yesss! Novel? yes Ethical? yes Relevant? Yes!! :)
```



4. What variables are relevant to your research question?

```{r}
# demographic metrics, and visibility within gievn research topic
```


5. Demonstrate how you would use ifelse() to make a given continuous variable into a dichtomous factor variable

```{r}
# if (year =< 1970){
#   print ('Modern Art Era')
# ifelse 
#   print('Contemporary Art Era')
# }
```





## NYC HANES Diabetes subset Questions

Data can be accessed at: https://raw.githubusercontent.com/fredlapolla/VilcekSummer26/refs/heads/main/NYC_HANES_DIAB.csv

6. Use apply() to find the median value of CADMIUM, LEAD CHOLESTEROLTOTAL, GLUCOSESI, CREATININESI, and CREATININE

```{r}
# apply(nyc[,11:18]), 2,median, na.rm = TRUE)
# answers : 
#   CADMIUM:0.2936  LEAD:1.056  CHOLTTL: 183.000  GLUC: 4.88  CREATININESI: 70.72 CREATININE: 0.800
```


7.  Do the same using sapply()

```{r}
# sapply(nyc[,11:19], median, na.rm = T)
# because it is sapply, i can take out the column bc it only runs on columns :)
```

8. Use a for loop to find the median value of columns 16:19

```{r}
#   for (i in 16:19){
#   value <- median(nyc[,i], na.rm = T)
#   print(c(colnames(nyc[i]), value))
#}
# ANSWERS:
#   CHOL: 183, GLUC: 4.88, CREATINNINES: 70.72, 
# CREATININE: 0.8
```


9. Make a for loop to find the standard deviation (hint sd() ) of CHOLESTEROLTOTAL, GLUCOSESI, CREATININESI, AND CREATININE

```{r}
#for (i in 16:19){
#   value <- sd(nyc[,i], na.rm = T)
#   print(c(colnames(nyc[i]), value))
# ANSWERS:
#   CHOL: 38.68, GLUC: 1.67, CREATINNINES: 26.51, 
# CREATININE: 0.29
```




10. Use ifelse to create a new variable of high and normal A1C dichotomizing with 7 as a threshold for high levels

```{r}
# nyc <- na.omit(nyc)
# nyc$HIGHA1C<- ifelse(nyc$A1C > 7,  1,  0)
# nyc$HIGHA1C <- factor(nyc$HiChol)
# summary(nyc$HIGHA1C)

# I DONT KNOW SRY 

```
